---
title: "Example spreadsheet data"
author: "Brian J. Knaus"
date: "`r format(Sys.time(), '%Y, %B %d')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## An example spreadsheet

In order to explore issues with using spreadsheet data we first need to manufacture an example.
Here I use the `iris` dataset to manufacture an example.
First we'll load the data set.

```{r}
data(iris)
head(iris)
class(iris)
str(iris)
```

## Add bad practices


To provide examples of bad practices we'll add them to the file.

```{r}
iris <- iris[,c(5, 1:4)]
iris$Species <- as.character(iris$Species)
iris$Species[10] <- "seto#sa" # Comment character
iris$Species[20] <- "setosa " # Trailing white space
iris$Species[30] <- "Setosa " # Mixed case
```



## Write to file

In the RMarkdown here I have set `eval=FALSE` so we do not clobber any existing files.
Set `eval=TRUE` or run the line manually if you would actually like to create the file.


```{r, eval=FALSE}
openxlsx::write.xlsx(iris, file = "iris_data.xlsx", sheetName = "iris")
```


We can also use a comma delimited file as follows.


```{r, eval=FALSE}
write.table(iris, "iris.csv", sep = ",", row.names = FALSE, quote = FALSE)
```


## Validate

Now we can attempt to read the file in to validate our changes.


```{r, eval=FALSE}
iris <- readxl::read_excel("iris_data.xlsx")
iris <- read.table("iris.csv", sep = ",")
```

