---
title: "Reading data"
author: "Brian J. Knaus"
date: "11/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Iris *.csv

I've created a file in this repository as an example that has known issues in it.
We use this as an example of how to troubleshoot problematic files.
We'll use `read.table()` to try to read it in.

```{r}
myIris <- read.table("iris.csv")
```

Or first attempt appears to be successful.
Let's explore what we read in to try and validate this.
Use the function `head()` to see what we read in.

```{r}
head(myIris)
str(myIris)
```

The function `read.table()` does not expect a header by default.
It also does not expect comma delimitation and instead expects white space delimiting.
If we set these parameters we can make another attempt to read the data in.

```{r}
myIris <- read.table("iris.csv", header = TRUE, sep = ",")
```


This time we should throw an error.
The error complains about line 10.
If we view the file in the RSudio we see the is a pound sign (#) on line 11.
R complains about line 10 because it is now interpreting the header as the first line.
The pound sign is interpretated as a comment character.
This means that R ignores everything on the row after the pound sign.
This creates a situation where this row has fewer elements than the previous lines and thus, we get an error.
If we fix this in a text editor or a spreadsheet we should now be able to read the data in.


```{r}
myIris <- read.table("iris.csv", header = TRUE, sep = ",")
head(myIris)
str(myIris)
```


## Changing class


When R reads the data in it tries to decide what type of data it is so it can use an appropriate container.
Its actually pretty good at this.
But sometimes we need to help it.
We can change the character `Sepal.Length` to a factor.


```{r}
myIris$Sepal.Length <- as.factor(myIris$Sepal.Length)
str(myIris)
```


We see that we now have a factor.
Converting this factor back presents a problem.


```{r}
as.numeric(myIris$Sepal.Length)
```


Instead of getting our numerics back we get a bunch of integers.
This is because R keeps track of the levels by numbering them (one through the number of levels we have).
This is why we get these integers.
In order to get our numbers back we need to first convert it to a character.

```{r}
myIris$Sepal.Length <- as.numeric(as.character(myIris$Sepal.Length))
str(myIris)
```

