---
title: "FASTA data"
author: "Brian J. Knaus"
date: "11/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read FASTA data in

DNA sequence information often come in the FASTA format.
We can use several functions in the R package `ape` to read this data into R.

```{r}
library(ape)
myDna <- read.dna("Phytopthora-ID-ITS.fasta", format = "fasta")
myDna
myDna[1]
print(myDna)
```

Note that our `DNAbin` object is storing the data as a `list`.
This is because the data is unaligned and different sequences are different lengths.


## Making system calls


When we anlyze data we typically want it to be aligned.
I do not know of any tool within R that will align sequences.
This is probably because it is computationally intensive.
But we can make system calls from within R to other command line tools.
Date is a simple example.


```{r}
system("date")
```

Compare this with the output you get from the terminal.
If you're using Windows, I'm not sure what this will do.

We can use the `maffft` aligner to align our sequences.

```{r, eval=FALSE}
system("mafft PhytoID.fa > PhytoID2.fa")
```


Now we can read our FASTA data into R as we did previously.


```{r}
myDna <- read.dna("Phytopthora-ID-ITS_aligned.fasta", format = "fasta")
myDna
```


Now our data is stored as a matrix.
The aligner creates insertions in the sequences so that they are all the same length.
Now we can manipulate the data in a manner similar to how we work with matrices.


```{r}
myDna[1:2, 1:10]
myDna[1,]
```


## A quick analysis


Now that we have a matrix we can perform a quick analysis.
First we'll create a distance matrix and use that to create a neighbor-joining tree.


```{r}
myDist <- dist.dna(myDna)
myNj <- nj(myDist)
plot(myNj)
```


